Online Resources 1: Supplementary results and tables

### 2025-06-19 analysis/supplementary-materials/ESM_1.qmd
Review corrections. Tables and graphs altered to include flakes and elongated blanks.

# Mutate level 5 to add phases: top of 5 (4E/5) and bottom of 5 (5), following Belmiro et al. (2020).
phase_dataset <- terrace_dataset |> 
  mutate(Level = case_when(
    Level == 5 & Z > 24.1 ~ "4E/5",
    Level == 5 & Z < 24.1 ~ "5",
    Level == "7C" | Level == 8 ~ "7", # Collapsing samples from Level 7C and Level 8 into 7 to homogenise dataset
    TRUE ~ Level)) |> # Adding technocomplex name based on the levels
  mutate(Technocomplex = case_when(
    Level == "4E" | Level == "4E/5" | Level == 5 ~ "Proto-Solutrean",
    Level == 6 | Level == 7 ~ "Gravettian")) |> 
  mutate(RETOUCHEDPIECETYPOLOGY = case_when(
    RETOUCHEDPIECETYPOLOGY == "Splintered piece with notch" | RETOUCHEDPIECETYPOLOGY == "Splintered piece with lateral retouch" | RETOUCHEDPIECETYPOLOGY == "Splintered piece fragment" ~ "Splintered piece",
    RETOUCHEDPIECETYPOLOGY == "Retouched blade" | RETOUCHEDPIECETYPOLOGY == "Retouched bladelet" | RETOUCHEDPIECETYPOLOGY == "Retouched flake" | RETOUCHEDPIECETYPOLOGY == "Retouched burin spall" ~ "Retouched blank",
    RETOUCHEDPIECETYPOLOGY == "Retouched bladelet frag" | RETOUCHEDPIECETYPOLOGY == "Retouched flake frag" | RETOUCHEDPIECETYPOLOGY == "Retouched flake siret fragment"  ~ "Retouched blank fragment",
    RETOUCHEDPIECETYPOLOGY == "Multiple burin" | RETOUCHEDPIECETYPOLOGY == "Burin splintered piece" | RETOUCHEDPIECETYPOLOGY == "Burin on truncation" | RETOUCHEDPIECETYPOLOGY == "Retouched burin" ~ "Burin",
    RETOUCHEDPIECETYPOLOGY == "Dentilhado" ~ "Denticulate",
    RETOUCHEDPIECETYPOLOGY == "Right truncation on retouched piece" ~ "Truncation",
    RETOUCHEDPIECETYPOLOGY == "Backed bladelet" | RETOUCHEDPIECETYPOLOGY == "Backed flake" | RETOUCHEDPIECETYPOLOGY == "Double backed bladelet" | RETOUCHEDPIECETYPOLOGY == "Backed bladelet parcial" ~ "Backed tool",
    RETOUCHEDPIECETYPOLOGY == "Carenated endscraper" ~ "Endscraper",
    TRUE ~ RETOUCHEDPIECETYPOLOGY))

@@

# Tables with class by raw material type (n and %)
general_table_som <- phase_dataset |> 
  dplyr::mutate(CLASS = case_when(
    CLASS == "BlankFrag" ~ "Blank fragment",
    CLASS == "BurinSpall" ~ "Burin spall",
    CLASS == "CoreFrag" ~ "Core fragment",
    CLASS == "CorePreparProd" ~ "Core prep product",
    CLASS == "RetouchedPiece" ~ "Retouched",
    CLASS == "RetouchedPieceFrag" ~ "Retouched fragment",
    TRUE ~ CLASS)) |> 
  dplyr::mutate(RMTYPE = case_when(
      RMTYPE == "T2" | RMTYPE == "T3" | RMTYPE == "T4" | RMTYPE == "T5" ~ "T2-5",
      TRUE ~ RMTYPE)) |> 
  filter(CLASS != "Chip")

## For Gravettian
gravettian_general_som <- general_table_som |>
  filter(Level %in% c(6, 7)) |>
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1","T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL", "INDET"))) |>
  tabyl(CLASS, RMTYPE) |> 
  adorn_totals(c("row", "col")) |> 
  adorn_percentages("col") |> 
  adorn_pct_formatting(2) |> 
  adorn_ns("front") |> 
  mutate(across(where(is.character), ~ gsub("^0   \\(0\\.00%\\)$", " ", .)))

## For Proto-Solutrean (4E and top of 5)
proto_general_som <- general_table_som |> 
  filter(Level %in% c("4E", "4E/5")) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1","T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL", "INDET"))) |>
  tabyl(CLASS, RMTYPE) |> 
  adorn_totals(c("row", "col")) |> 
  adorn_percentages("col") |> 
  adorn_pct_formatting(2) |> 
  adorn_ns("front") |> 
  mutate(across(where(is.character), ~ gsub("^0   \\(0\\.00%\\)$", " ", .)))

## For Solutrean
solutrean_general_som <- solutrean_dataset |> 
  dplyr::mutate(RMTYPE = case_when(
      RMTYPE == "T2" | RMTYPE == "T3" | RMTYPE == "T4" | RMTYPE == "T5" ~ "T2-5",
      TRUE ~ RMTYPE)) |>
  dplyr::mutate(CLASS = case_when(
    CLASS == "ElongBlank" | CLASS == "Flake" ~ "Blank",
    CLASS == "ElongBlankFrag" | CLASS == "FlakeFrag" ~ "Blank fragment",
    CLASS == "Thinning flake" ~ "Blank",
    CLASS == "CorePrepProd" ~ "Core prep product",
    CLASS == "RetouchedPiece" | CLASS == "Retouched piece" ~ "Retouched",
    CLASS == "Retouched piece fragment" ~ "Retouched fragment",
    TRUE ~ CLASS)) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1","T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "T11", "TL", "INDET"))) |>
  tabyl(CLASS, RMTYPE) |> 
  adorn_totals(c("row", "col")) |> 
  adorn_percentages("col") |> 
  adorn_pct_formatting(2) |> 
  adorn_ns("front") |> 
  mutate(across(where(is.character), ~ gsub("^0   \\(0\\.00%\\)$", " ", .)))

@@

# Solutrean statistics
# DS preparation
solutrean_stats <- solutrean_dataset |> 
  filter(RMTYPE != "INDET") |> 
  mutate(CLASS = case_when(
    CLASS == "ElongBlank" | CLASS == "Flake" ~ "Blank",
    TRUE ~ CLASS)) |> 
  dplyr::mutate(RMTYPE = case_when(
      RMTYPE == "T2" | RMTYPE == "T3" | RMTYPE == "T4" | RMTYPE == "T5" ~ "T2-5",
      TRUE ~ RMTYPE))

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
solutrean_stats_blank <- solutrean_stats |>  
  filter(CLASS == "Blank") |> 
  select(RMTYPE, CLASS, MAXWIDTH, LENGTH, WEIGHT) |> 
  drop_na() |> 
  group_by(RMTYPE, CLASS) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "T11", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))
# For cores
solutrean_stats_cores <- solutrean_stats |> 
  filter(CLASS == "Core") |>
  select(RMTYPE, CLASS, MAXWIDTH, LENGTH, WEIGHT) |> 
  drop_na() |> 
  group_by(RMTYPE, CLASS) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "T11", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))

@@

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
proto_stats_blank <- proto_stats |> 
  filter(CLASS == "Blank") |> 
  select(RMTYPE, CLASS, MAXWIDTH, LENGTH, WEIGHT) |> 
  group_by(RMTYPE, CLASS) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))
# For cores
proto_stats_cores <- proto_stats |> 
  filter(CLASS == "Core") |> 
  select(RMTYPE, CLASS, MAXWIDTH, LENGTH, WEIGHT) |> 
  group_by(RMTYPE, CLASS) |>
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))

@@

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
gravettian_stats_blank <- gravettian_stats |> 
  filter(CLASS == "Blank") |> 
  select(RMTYPE, CLASS, MAXWIDTH, LENGTH, WEIGHT) |> 
  group_by(RMTYPE, CLASS) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))
# For cores
gravettian_stats_cores <- gravettian_stats |> 
  filter(CLASS == "Core") |> 
  select(RMTYPE, CLASS, MAXWIDTH, LENGTH, WEIGHT) |> 
  group_by(RMTYPE, CLASS) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))

@@

```{r}
#| label: tbl-gravettian
#| tbl-cap: "Mean and standard deviation (SD) for Max width, Length and Weight measurements of blanks (a) and cores (b) by chert type of the Gravettian assemblage (levels 6-7) of the Terrace area."
#| tbl-subcap: 
#|  - "Mean and SD for Max width, Length and Weight measurements of blanks."
#|  - "Mean and SD for Max width, Length and Weight measurements of cores."
#| layout-nrow: 2

kable(gravettian_stats_blank)
kable(gravettian_stats_cores)

@@

```{r}
#| label: tbl-proto
#| tbl-cap: "Mean and standard deviation (SD) for Max width, Length and Weight measurements of blanks (a) and cores (b) by chert type of the Proto-Solutrean assemblage (levels 4E and top of level 5) of the Terrace area."
#| tbl-subcap: 
#|  - "Mean and SD for Max width, Length and Weight measurements of blanks."
#|  - "Mean and SD for Max width, Length and Weight measurements of cores."
#| layout-nrow: 2

kable(proto_stats_blank)
kable(proto_stats_cores)

@@

```{r}
#| label: tbl-solutrean
#| tbl-cap: "Mean and standard deviation (SD) for Max width, Length and Weight measurements of blanks (a) and cores (b) by chert type of the Solutrean assemblage (levels A to C) of the Shelter area."
#| tbl-subcap: 
#|  - "Mean and SD for Max width, Length and Weight measurements of blanks."
#|  - "Mean and SD for Max width, Length and Weight measurements of cores."
#| layout-nrow: 2

kable(solutrean_stats_blank)
kable(solutrean_stats_cores)

>>>>

# Mutate level 5 to add phases: top of 5 (4E/5) and bottom of 5 (5), following Belmiro et al. (2020).
phase_dataset <- terrace_dataset |> 
  mutate(Level = case_when(
    Level == 5 & Z > 24.1 ~ "4E/5",
    Level == 5 & Z < 24.1 ~ "5",
    Level == "7C" | Level == 8 ~ "7", # Collapsing samples from Level 7C and Level 8 into 7 to homogenise dataset
    TRUE ~ Level)) |> # Adding technocomplex name based on the levels
  mutate(Technocomplex = case_when(
    Level == "4E" | Level == "4E/5" | Level == 5 ~ "Proto-Solutrean",
    Level == 6 | Level == 7 ~ "Gravettian")) |> 
  mutate(RETOUCHEDPIECETYPOLOGY = case_when(
    RETOUCHEDPIECETYPOLOGY == "Splintered piece with notch" | RETOUCHEDPIECETYPOLOGY == "Splintered piece with lateral retouch" | RETOUCHEDPIECETYPOLOGY == "Splintered piece fragment" ~ "Splintered piece",
    RETOUCHEDPIECETYPOLOGY == "Retouched blade" | RETOUCHEDPIECETYPOLOGY == "Retouched bladelet" | RETOUCHEDPIECETYPOLOGY == "Retouched flake" | RETOUCHEDPIECETYPOLOGY == "Retouched burin spall" ~ "Retouched blank",
    RETOUCHEDPIECETYPOLOGY == "Retouched bladelet frag" | RETOUCHEDPIECETYPOLOGY == "Retouched flake frag" | RETOUCHEDPIECETYPOLOGY == "Retouched flake siret fragment"  ~ "Retouched blank fragment",
    RETOUCHEDPIECETYPOLOGY == "Multiple burin" | RETOUCHEDPIECETYPOLOGY == "Burin splintered piece" | RETOUCHEDPIECETYPOLOGY == "Burin on truncation" | RETOUCHEDPIECETYPOLOGY == "Retouched burin" ~ "Burin",
    RETOUCHEDPIECETYPOLOGY == "Dentilhado" ~ "Denticulate",
    RETOUCHEDPIECETYPOLOGY == "Right truncation on retouched piece" ~ "Truncation",
    RETOUCHEDPIECETYPOLOGY == "Backed bladelet" | RETOUCHEDPIECETYPOLOGY == "Backed flake" | RETOUCHEDPIECETYPOLOGY == "Double backed bladelet" | RETOUCHEDPIECETYPOLOGY == "Backed bladelet parcial" ~ "Backed tool",
    RETOUCHEDPIECETYPOLOGY == "Carenated endscraper" ~ "Endscraper",
    TRUE ~ RETOUCHEDPIECETYPOLOGY)) |> 
  # --Article review--
  mutate(elong = MAXWIDTH*2-LENGTH) %>%
  mutate(CLASS_ELONG = case_when( 
    CLASS == "Blank" & elong > 0 ~ "Flake",
    CLASS == "Blank" & elong <= 0 ~ "ElongatedProd",
    TRUE ~ CLASS))

@@

# Tables with class by raw material type (n and %)
general_table_som <- phase_dataset |> 
  dplyr::mutate(CLASS_ELONG = case_when(
    CLASS_ELONG == "BlankFrag" ~ "Blank fragment",
    CLASS_ELONG == "ElongatedProd" ~ "Elongated blank",
    CLASS_ELONG == "BurinSpall" ~ "Burin spall",
    CLASS_ELONG == "CoreFrag" ~ "Core fragment",
    CLASS_ELONG == "CorePreparProd" ~ "Core prep product",
    CLASS_ELONG == "RetouchedPiece" ~ "Retouched",
    CLASS_ELONG == "RetouchedPieceFrag" ~ "Retouched fragment",
    TRUE ~ CLASS_ELONG)) |> 
  dplyr::mutate(RMTYPE = case_when(
      RMTYPE == "T2" | RMTYPE == "T3" | RMTYPE == "T4" | RMTYPE == "T5" ~ "T2-5",
      TRUE ~ RMTYPE)) |> 
  filter(CLASS != "Chip")

## For Gravettian
gravettian_general_som <- general_table_som |>
  filter(Level %in% c(6, 7)) |>
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1","T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL", "INDET"))) |>
  mutate(CLASS_ELONG = factor(CLASS_ELONG, levels = c("Flake", "Elongated blank", "Blank fragment", "Burin spall", "Core", "Core fragment", "Core prep product", "Retouched", "Retouched fragment", "Shatter"))) |> 
  tabyl(CLASS_ELONG, RMTYPE) |> 
  adorn_totals(c("row", "col")) |> 
  adorn_percentages("col") |> 
  adorn_pct_formatting(2) |> 
  adorn_ns("front") |> 
  mutate(across(where(is.character), ~ gsub("^0   \\(0\\.00%\\)$", " ", .)))

## For Proto-Solutrean (4E and top of 5)
proto_general_som <- general_table_som |> 
  filter(Level %in% c("4E", "4E/5")) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1","T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL", "INDET"))) |>
  mutate(CLASS_ELONG = factor(CLASS_ELONG, levels = c("Flake", "Elongated blank", "Blank fragment", "Burin spall", "Core", "Core fragment", "Core prep product", "Retouched", "Retouched fragment", "Shatter"))) |> 
  tabyl(CLASS_ELONG, RMTYPE) |> 
  adorn_totals(c("row", "col")) |> 
  adorn_percentages("col") |> 
  adorn_pct_formatting(2) |> 
  adorn_ns("front") |> 
  mutate(across(where(is.character), ~ gsub("^0   \\(0\\.00%\\)$", " ", .)))

## For Solutrean
solutrean_general_som <- solutrean_dataset |> 
  dplyr::mutate(RMTYPE = case_when(
      RMTYPE == "T2" | RMTYPE == "T3" | RMTYPE == "T4" | RMTYPE == "T5" ~ "T2-5",
      TRUE ~ RMTYPE)) |>
  dplyr::mutate(CLASS = case_when(
    CLASS == "ElongBlank" ~ "Elongated blank",
    CLASS == "ElongBlankFrag" | CLASS == "FlakeFrag" ~ "Blank fragment",
    CLASS == "CorePrepProd" ~ "Core prep product",
    CLASS == "RetouchedPiece" | CLASS == "Retouched piece" ~ "Retouched",
    CLASS == "Retouched piece fragment" ~ "Retouched fragment",
    TRUE ~ CLASS)) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1","T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "T11", "TL", "INDET"))) |>
    mutate(CLASS = factor(CLASS, levels = c("Flake", "Elongated blank", "Thinning flake", "Blank fragment", "Burin spall", "Core", "Core fragment", "Core prep product", "Retouched", "Retouched fragment", "Shatter"))) |> 
  tabyl(CLASS, RMTYPE) |> 
  adorn_totals(c("row", "col")) |> 
  adorn_percentages("col") |> 
  adorn_pct_formatting(2) |> 
  adorn_ns("front") |> 
  mutate(across(where(is.character), ~ gsub("^0   \\(0\\.00%\\)$", " ", .)))

@@

# Solutrean statistics
# DS preparation
solutrean_stats <- solutrean_dataset |> 
  filter(RMTYPE != "INDET") |> 
  mutate(CLASS = case_when(
    CLASS == "ElongBlank" ~ "Elongated blank",
    TRUE ~ CLASS)) |> 
  dplyr::mutate(RMTYPE = case_when(
      RMTYPE == "T2" | RMTYPE == "T3" | RMTYPE == "T4" | RMTYPE == "T5" ~ "T2-5",
      TRUE ~ RMTYPE))

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
solutrean_stats_blank <- solutrean_stats |>  
  filter(CLASS == "Flake") |> 
  select(RMTYPE, CLASS, MAXWIDTH, LENGTH, WEIGHT) |> 
  drop_na() |> 
  group_by(RMTYPE, CLASS) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "T11", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))
# For elongated blanks (complete) --Article review--
solutrean_stats_elong <- solutrean_stats |>  
  filter(CLASS == "Elongated blank") |> 
  select(RMTYPE, CLASS, MAXWIDTH, LENGTH, WEIGHT) |> 
  drop_na() |> 
  group_by(RMTYPE, CLASS) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "T11", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))
# For cores
solutrean_stats_cores <- solutrean_stats |> 
  filter(CLASS == "Core") |>
  select(RMTYPE, CLASS, MAXWIDTH, LENGTH, WEIGHT) |> 
  drop_na() |> 
  group_by(RMTYPE, CLASS) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "T11", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))

@@

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
proto_stats_blank <- proto_stats |> 
  filter(CLASS_ELONG == "Flake") |> 
  select(RMTYPE, CLASS_ELONG, MAXWIDTH, LENGTH, WEIGHT) |> 
  group_by(RMTYPE, CLASS_ELONG) |> 
  rename(CLASS = CLASS_ELONG) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))
# For elongated blanks (complete) --Article review--
proto_stats_elong <- proto_stats |> 
  dplyr::mutate(CLASS_ELONG = case_when(
    CLASS_ELONG == "ElongatedProd" ~ "Elongated blank",
    TRUE ~ CLASS_ELONG)) |> 
  filter(CLASS_ELONG == "Elongated blank") |> 
  select(RMTYPE, CLASS_ELONG, MAXWIDTH, LENGTH, WEIGHT) |> 
  group_by(RMTYPE, CLASS_ELONG) |> 
  rename(CLASS = CLASS_ELONG) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))
# For cores
proto_stats_cores <- proto_stats |> 
  filter(CLASS == "Core") |> 
  select(RMTYPE, CLASS, MAXWIDTH, LENGTH, WEIGHT) |> 
  group_by(RMTYPE, CLASS) |>
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))

@@

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
gravettian_stats_blank <- gravettian_stats |> 
  filter(CLASS_ELONG == "Flake") |> 
  select(RMTYPE, CLASS_ELONG, MAXWIDTH, LENGTH, WEIGHT) |> 
  group_by(RMTYPE, CLASS_ELONG) |> 
  rename(CLASS = CLASS_ELONG) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))
# For elongated blanks (complete) --Article review--
gravettian_stats_elong <- gravettian_stats |> 
  dplyr::mutate(CLASS_ELONG = case_when(
    CLASS_ELONG == "ElongatedProd" ~ "Elongated blank",
    TRUE ~ CLASS_ELONG)) |> 
  filter(CLASS_ELONG == "Elongated blank") |> 
  select(RMTYPE, CLASS_ELONG, MAXWIDTH, LENGTH, WEIGHT) |> 
  group_by(RMTYPE, CLASS_ELONG) |> 
  rename(CLASS = CLASS_ELONG) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))
# For cores
gravettian_stats_cores <- gravettian_stats |> 
  filter(CLASS == "Core") |> 
  select(RMTYPE, CLASS, MAXWIDTH, LENGTH, WEIGHT) |> 
  group_by(RMTYPE, CLASS) |> 
  mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2-5", "T6","T7", 
                                        "T8", "T9", "T10", "TL"))) |>
  summarise(across(c(MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd)))

@@

```{r}
#| label: tbl-gravettian
#| tbl-cap: "Mean and standard deviation (SD) for Max width, Length and Weight measurements of flakes (a), elongated blanks (b) and cores (c) by chert type of the Gravettian assemblage (levels 6-7) of the Terrace area."
#| tbl-subcap: 
#|  - "Mean and SD for Max width, Length and Weight measurements of flakes."
#|  - "Mean and SD for Max width, Length and Weight measurements of elongated blanks."
#|  - "Mean and SD for Max width, Length and Weight measurements of cores."
#| layout-nrow: 3

kable(gravettian_stats_blank)
kable(gravettian_stats_elong)
kable(gravettian_stats_cores)

@@

```{r}
#| label: tbl-proto
#| tbl-cap: "Mean and standard deviation (SD) for Max width, Length and Weight measurements of flakes (a), elongated blanks (b) and cores (c) by chert type of the Proto-Solutrean assemblage (levels 4E and top of level 5) of the Terrace area."
#| tbl-subcap: 
#|  - "Mean and SD for Max width, Length and Weight measurements of flakes."
#|  - "Mean and SD for Max width, Length and Weight measurements of elongated blanks."
#|  - "Mean and SD for Max width, Length and Weight measurements of cores."
#| layout-nrow: 3

kable(proto_stats_blank)
kable(proto_stats_elong)
kable(proto_stats_cores)

@@

```{r}
#| label: tbl-solutrean
#| tbl-cap: "Mean and standard deviation (SD) for Max width, Length and Weight measurements of flakes (a), elongated blanks (b) and cores (c) by chert type of the Solutrean assemblage (levels A to C) of the Shelter area."
#| tbl-subcap: 
#|  - "Mean and SD for Max width, Length and Weight measurements of flakes."
#|  - "Mean and SD for Max width, Length and Weight measurements of elongated blanks."
#|  - "Mean and SD for Max width, Length and Weight measurements of cores."
#| layout-nrow: 3

kable(solutrean_stats_blank)
kable(solutrean_stats_elong)
kable(solutrean_stats_cores)
